---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
<<<<<<< Updated upstream


## What is mean total number of steps taken per day?


=======
```{r}
library(data.table)
data <- read.table("activity.csv", header=T, sep=",")
```

## What is mean total number of steps taken per day?
```{r, echo=TRUE}
library(sqldf) ### SQL package based on SQLite
# Make a histogram of the total number of steps taken each day
data1 <- sqldf("select date,sum(steps) as total_steps from data where steps is not null group by 1")
hist(data1$total_steps, col="red", xlab="Total Steps", main="Total Steps per Day")
# Calculate and report the mean and median total number of steps taken per day
MeanSteps <- mean(data1$total_steps)
print(MeanSteps)
MedianSteps <- median(data1$total_steps)
print(MedianSteps)
```

## What is the average daily activity pattern?
```{r, echo=TRUE}
# Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
data2 <- sqldf("select interval as intervalKey, avg(steps) as averageSteps from data where steps is not null group by 1")
library(ggplot2)
ggplot(data2, aes(x=intervalKey)) + geom_line(aes(y=averageSteps)) + labs(y="Average # of Steps", x="Interval") + ggtitle("Avg. Daily Steps")
# Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
## first we find the max value
maxSteps <- max(data2$averageSteps)
## then the max interval
maxInterval <- data2[data2$averageSteps == maxSteps,]
print(maxInterval$interval)
```
## Imputing missing values
```{r, echo=TRUE}
# Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
missing_values <- sqldf("select count(*) as steps from data where steps is null")
print (missing_values)
# Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc. and Create a new dataset that is equal to the original dataset but with the missing data filled in.
data$steps <- as.numeric(data$steps) # I had to change the datatype to numeric
data3 <- sqldf("select data.date,data.interval, case when steps is null then averageSteps else steps end as steps from data inner join data2 on data.interval=data2.intervalKey")
# Make a histogram of the total number of steps taken each day
data4 <- sqldf("select date,sum(steps) as totalSteps from data3 where steps is not null group by 1")
hist(data4$totalSteps, col="blue", xlab="Total Steps", main="Total Steps per Day")
# and Calculate and report the mean and median total number of steps taken per day
meanSteps_new <- mean(data4$totalSteps)
print(meanSteps_new)
medianSteps_new <- median(data4$totalSteps)
print(medianSteps_new)
```
Q: What is the impact of imputing missing data on the estimates of the total daily number of steps?

A: mean stays the same and median increases by one.

## Are there differences in activity patterns between weekdays and weekends?
```{r, echo=TRUE}
#Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
data5 <- sqldf("select case when strftime('%w', date) in ('0','6') then 'WeekEnd' else 'WeekDay' end as dayType, interval, avg(steps) as avgSteps from data3 group by 1,2") # using strftime with %w for day of week (0 is Sunday 1 is Monday)
# Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
ggplot(data5, aes(x=interval)) + geom_line(aes(y=avgSteps), col="blue") + facet_wrap(~dayType, ncol=1) +
labs(y="Average # of Steps", x="Interval")
```

```{text}
* All figures are also available in figuers directory
```


